---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load the FAS_litters data

```{r}
litters_df = read_csv("./data/FAS_litters.csv", na = c("NA", ".", ""))
litters_df = janitor :: clean_names(litters_df)
litters_df
```

## 'select'

Chose some columns and not others.

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
select(litters_df, -litter_number)
```

Renaming columns

```{r}
select(litters_df, GROUP = group, LITTer_NuMBer = litter_number)
```

can also use rename if you want to rename a few without getting rid of the rest of the columns

```{r}
rename(litters_df, GROUP = group, LITTer_NuMBer = litter_number)
```

select helpers

```{r}
select(litters_df, starts_with("gd"))
```

moves the column you want up to the front
```{r}
select(litters_df, litter_number, everything())
```

relocate a column to the beginning or other places
```{r}
relocate(litters_df, litter_number)
```

basically there are a bunch of helper functions that do things that you can also do with select.

## 'filter'

while select operates on columns, filter operates on rows. filter needs conditions (true/false)

```{r}
filter(litters_df, gd0_weight >= 22)
```

use the double equal sign when posing the question , rather than telling it to set things equal to 20

```{r}
filter(litters_df, gd_of_birth == 20)
```

adding the exclamation point outside of the parenthesis does the opposite, gives you everything where its not equal to 20. we also added a second way to do this below
```{r}
filter(litters_df, !(gd_of_birth == 20))
filter(litters_df, gd_of_birth != 20)
```

```{r}
filter(litters_df, gd0_weight >=22, gd_of_birth==20)
```

```{r}
filter(litters_df, group == "Mod8")
```

this says where group is either of these two options
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```

## 'mutate'
```{r}
litters_df
```

```{r}
drop_na(litters_df)
mutate(litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

## 'arrange'

```{r}
arrange(litters_df, pups_born_alive)
```

## '%>%'

```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv", na = c("NA", ".", ""))
litters_clean_name = janitor :: clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)
litters_mutated

litters_without_missing = drop_na(litters_mutated, gd0_weight)
litters_without_missing
```

USE the PIPE OPERATOR INSTEAD

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv", na = c("NA", ".", "")) %>% 
  janitor :: clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)
litters_df
```



